---
title: "Reproducible Research Project 1"
author: "Sylvia Badon"
date: "11/3/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and Preprocessing the Data

```{r}
d <- read.csv("I:/K99R00/Training/Accelerometry/R/Coursera Reproducible Research Course/Project 1/activity.csv")
d$date <- as.Date(d$date)
```


## What is the mean total number of steps taken per day? 

First, calculate the total number of steps per day, ignoring missing values.

```{r}
total <- tapply(d$steps, d$date, sum, na.rm=TRUE)
```

Then plot total steps per day in a histogram.

```{r, echo=TRUE}
hist(total, main="Histogram of total steps per day", xlab="Total steps in one day")
```


Finally, calculate the mean and median of the total number of steps per day. 

```{r}
mean.steps <- as.integer(mean(total))
median.steps <- median(total)
```

The mean total steps per day was `r mean.steps`. The median total steps per day was `r median.steps`.


## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval and average number of steps taken, averaged across all days, ignoring missing values.

```{r}
by.int <- tapply(d$steps, d$interval, mean, na.rm=TRUE)
by.int2 <- cbind(rownames(by.int), as.data.frame(by.int))
colnames(by.int2) <- c("interval", "avg.steps")

plot(by.int2$interval, by.int2$avg.steps, type="l", main="Average steps per 5-minute interval", xlab="5-minute interval", ylab="Average number of steps")

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` {r}
max.int <- by.int2[by.int2$avg.steps==max(by.int2$avg.steps), 1]
```

The `r max.int` interval contains the maximum number of average steps across all the days in the dataset.


## Imputing Missing Values

Calculate and report total number of missing values in the dataset.

``` {r}
comp <- complete.cases(d)
comp.case <- as.list(table(comp))
num.miss <- comp.case$'FALSE'
```

The total number of rows with missing values in the dataset is `r num.miss`. 

Fill in missing values with mean of that 5-min interval.

```{r}
with.intavg <- merge(d, by.int2, by="interval")
with.intavg$steps <- ifelse(is.na(with.intavg$steps), with.intavg$avg.steps, with.intavg$steps)
```

Make a histogram of the total number of steps taken per day using the imputed data.

```{r}
total2 <- tapply(with.intavg$steps, with.intavg$date, sum)
hist(total2, main="Histogram of total steps per day (with imputation)", xlab="Total steps in one day")
```

Calculate the mean and median of the total number of steps per day using the imputed dataset. 

```{r}
mean.steps2 <- as.integer(mean(total2))
median.steps2 <- as.integer(median(total2))
```

After imputation, the mean total steps per day was `r mean.steps2`. The median total steps per day was `r median.steps2`.


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable with two levels: weekday and weekend in the imputed dataset.

```{r}
with.intavg$day <- weekdays(with.intavg$date)
wkdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
with.intavg$day2 <- factor(ifelse(with.intavg$day %in% wkdays, "weekday", "weekend"))
```

Make a time series plot of the 5-minute interval and average number of steps taken, averaged across all weekdays or weekend days.

```{r}
weekday.data <- with.intavg[with.intavg$day2=="weekday" , ]
by.int_wkday <- tapply(weekday.data$steps, weekday.data$interval, mean, na.rm=TRUE)
by.int2_wkday <- cbind(rownames(by.int_wkday), as.data.frame(by.int_wkday))
colnames(by.int2_wkday) <- c("interval", "avg.steps")

weekend.data <- with.intavg[with.intavg$day2=="weekend" , ]
by.int_wkend <- tapply(weekend.data$steps, weekend.data$interval, mean, na.rm=TRUE)
by.int2_wkend <- cbind(rownames(by.int_wkend), as.data.frame(by.int_wkend))
colnames(by.int2_wkend) <- c("interval", "avg.steps")

par(mar=c(4.0, 4.1, 4.1, 0.1), mfrow=c(2,1))
plot(by.int2_wkend$interval, by.int2_wkend$avg.steps, type="l", main="Weekends", xlab=" ", ylab="Avg number of steps")
plot(by.int2_wkday$interval, by.int2_wkday$avg.steps, type="l", main="Weekdays", xlab="5-minute interval", ylab="Avg number of steps")
```
